### 2009-11-25 doc/manuals/janitor/tex_concept/concept.tex
Michael's bits.
Series of updates, need to verify the janitor webservice, still. Please merge.

\chapter{Programming concept}

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog's RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

@@

\section{States of runtime environments}
A major motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

@@

Upon presentation of a the package name to a Catalog, from which details about
the package are retrieved, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

\section{Janitor file system permissions}

In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

It is suggested to create a user and a group "janitor" for the Janitor.\\

@@

>>>>

\chapter{Technical Motivation and Concepts}

The current implementation, using the rather advanced concepts of the
semantic web, may seem unexpected. They were chosen since they are expected
to scale with all the extra demands forseen for the Janitor.

\section{Implementation}

\subsection{Choice of Perl}

 The main language for the implementation of the functionality of the
 dRTE's functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor.
 The language was perceived a side-issue, the complexity of the data
 expected was of major concern.

\subsection{Modular structure}

 In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog's RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

@@

\section{Concepts}

In the following, some paragraphs fail render it difficult to 
clearly distinguish between conceptional truths and the state
of the current implementation. Please read carefully.

\subsection{States of runtime environments}

A major motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production. It would be nice to
see the Janitor and/or the Catalog prepare for reviews by selected
users. This has not yet been implemented.

With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

@@

\subsection{Subset of that functionality as implemented today}
Upon presentation of a the package name to a Catalog, from which details about
the package are retrieved, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

%FIXME: needs further communication with Anders
%\section{Janitor file system permissions}
%
%In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
%a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
%i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
%
%Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
%the wrapper.\\
%
%It is suggested to create a user and a group "janitor" for the Janitor.\\

@@

\subsection{What happens during installation}

Register
\begin{enumerate}
 \item Look up Catalog and find corresponding runtime environments.
 \item Check if dependcies are available (installed or installable)
 \item Store job in \textit{registrationdir}
\end{enumerate}

Deploy
\begin{enumerate}
 \item Load job out of \textit{registrationdir}
 \item Look up Catalog and find corresponding runtime environments.
 \item Check if dependcies are available (installed or installable)
 \item Download RTEs into the \textit{downloaddir}
 \item $\lbrack$ to be continued or skipped $\rbrack$
\end{enumerate}

### 2009-11-25 doc/manuals/janitor/tex_future/future.tex
Series of updates to the Janitor documentation.
Michael's comments on future developments on the Janitor.
Series of updates, need to verify the janitor webservice, still. Please merge.

\section{Representation of dynamic REs in the information model}
Dynamic REs require an extended representation in the
information model.  The Application Software description should be able
to distinguish installed REs from installable REs, potentially offer
description of extended RE state-like information. This work is planned
to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
GLUE: https://forge.gridforum.org/sf/projects/glue-wg}.

@@

Future development of ARC aims at integrating grid computing with
workflow tools for the web services that have a growing user base in
bioinformatics. The challenge is to prepare REs for
programs or databases and to offer such concisely to users of the
workflow environments.  In the bioinformatics community, such are
today offered as web services.  This anticipated development instead
fosters the dynamic installation on the grid whenever appropriate to
allow for special computational demands in high-throughput analyses.
Conversely, because of the increased complexity of workflows with respect
to the already today not manually manageable number of REs, without an
automatism for the automated installation of software packages on the
grid, the use of workflows in grid computing seems mute.

@@

The RDF schema nicely prepares for the upcoming virtualisation of worker
nodes.  How exactly the dynamics are integrated will depend on how
dynamic the virtualisation of the nodes is. In the simplest scenario,
a worker node's CPU will only be occupied by a single virtual machine
and that will not be changed. In this case, there is no difference to
the setup of the Janitor with today's static setups.

@@

TODO:\\
\begin{itemize}
 \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
 \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
 \item Verfication
\end{itemize}

>>>>

\section{Representation of dynamic RTEs in the information model}
Dynamic RTEs require an extended representation in the
information model.  The Application Software description should be able
to distinguish installed RTEs from installable RTEs, potentially offer
description of extended RTE state-like information. This work is planned
to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
GLUE: https://forge.gridforum.org/sf/projects/glue-wg}.

@@

Future development of ARC aims at integrating grid computing with
workflow tools for the web services that have a growing user base in
bioinformatics. The challenge is to prepare RTEs for
programs or databases and to offer such concisely to users of the
workflow environments.  In the bioinformatics community, such are
today offered as web services.  This anticipated development instead
fosters the dynamic installation on the grid whenever appropriate to
allow for special computational demands in high-throughput analyses.
Conversely, because of the increased complexity of workflows with respect
to the already today not manually manageable number of RTEs, without an
automatism for the automated installation of software packages on the
grid, the use of workflows in grid computing seems mute.

@@

The RDF schema nicely prepares for the upcoming virtualisation of worker
nodes. Hereto, the {\tt BaseSystem} indicates a virtual image to which
further packages, the dynamic REs, would then be added.
How exactly the dynamics are integrated will depend on how
dynamic the virtualisation of the nodes is. In the simplest scenario,
a worker node's CPU will only be occupied by a single virtual machine
and that will not be changed. In this case, there is no difference to
the setup of the Janitor with today's static setups.

@@

%TODO:\\
%\begin{itemize}
% \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
% \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
% \item Verfication
%\end{itemize}

%(1) The HostInfoJanitor.pm queries the catalog for available RTE names.
%    It is fast, such that I assume a fork or a cache is currently not needed. 
%    > Nevertheless I installed an alarm of 15 Seconds

%(2) In order to enable the Grid-Manager to invoke Janitor, a wrapper rJanitor.c has
%    (already) been written (by Daniel). 
%    This wrapper gives Janitor the necessary rights using a suid bit. 
%    >It is installed and used but not yet tested well
%    >In case Daniel is right, it will work

% (3) > ... ?? hmm... well seems there is some information loss here... point 3... suppose it must have been important.

%(4) The function "check" (nowadays ``info'') was meant to return the information about a runtime 
%      environment. It may possible to merge with the deploy functionality 
%      if desired. (The Janitor source code will remain the same, only the interface will 
%      be changed.)
%      >Since this is mere beautification, this is postponed

%(5) The information about the runtime environments is passed to the backend
%    script. (It is not specifed yet how this has to be done)
%       o Installation directories
%       o Script to set environment variables
%    > Currently a ``deploy'' action will set symbolic links for *ALL* dynamically installed
%    > METAPACKAGES!  That means, nothing has to be done here. The backends script won't (hopefully)
%    > except the symbolic links and execute the corresponding script delievered along with the tarball 
%    > package. 
%    > The symbolic link will be removed respictvely on ``remove''
%
%    > TODO: Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change *that*
%
%    > TODO: Think about the user rights of the symbolic link...

%(6) Once the grid job has been finished, the token on the runtime environment will
%      be removed. (If no tokens left, the environment may be deinstalled)
%    > Well, that's nothing new for me, but maybe nice to know :-)

%(7) GLUE2 specification: Balazs like to have an interface for Janitor. This task is 
%   put on the agenda. It is still unclear how the interaction with that interface shall be 
%   realised. Janitor can only access data concerning the class ApplicationEnvironment.
%   Shall a XML snippet be returned?
%  > Alright now, almost 1.5 month later I know what he asked for.
%  >
%  > Please examine following files in src/services/a-rex/infoproviders/:
%    o ARC1ClusterInfo.pm:~744
%    o HostInfo.pm:~218
%    o HostInfoJanitor.pm
%  >
%  > The HostInfo.pm collects a list of manually and dynamically (HostInfoJanitor.pm) installed RTE *NAMES*.
%  > ARC1ClusterInfo.pm adds some hard coded information (description, states, etc.) to it and puts it into the schema.
%  >
%  > WHAT HAS TO BE DONE?
%  >
%  > HostInfo.pm has to create a *HASH MAP* with must have as much information as possible (description, states, etc.).
%  > Therefore HostInfoJanitor.pm has to be extended to gather more information out of the catalog.
%  > ARC1ClusterInfo.pm needs to be stultified such that the *HASH MAP* received from HostInfo.pm will simply mapped
%  > to the schema.

%(8) Verfication
%    o Of validity of tar-package (Store the catalog somewhere safe! Hash-value database?)
%    o Of the executer of Janitor (This can be done using a WEB SERVICE!!! )
%    o TODO: I experienced some problems in case some shell tools are not installed on clusters
%            i.e.  unzip
%            There should be some kind of check or RESTRICTION to what can be used

%(9) The missing feature for dependencies 
%    Assume, one wants to install package A which depends on package B.
%    Both are getting installed dynamically.
%    Right before job execution, the runtime scripts of package A is executed. 
%    Still missing is the execution of the runtime script of B!
%
%    Simple workaround: Request package B in job description too
%    Correct solution:  Dependencies have to be returned by Janitor and added to the job.description file!!!!!
%                       (Janitor info 21 or *EXTEND* Janitor deploy + Get Janitor output in states.cpp and put
%			it into the description file)

### 2009-11-25 doc/manuals/janitor/tex_maintenance/maintenance.tex
Series of updates, need to verify the janitor webservice, still. Please merge.

>>>>

\subsection{Automated transformation of install directory to dRTE}

The script 'prepareDRE.pl' was created to help with the transformation
of a readily installed software into a dynamic runtime environment.
I also prepares a complete catalog file that can be offered
individually or next to other catalogs. See the associated man page
prepareDRE(8) for details.
